---
title: "Automated E-Mail Updates Of CRAN Downloads Using Raspberry Pi 4"
description: |
  Want to have regular updates on how your R packages are doing on CRAN when it comes to downloads? This is one way using a Raspberry Pi 4 and e-mails.
author:
  - name: Yannik
    url: yannikbuhl.de
date: 2022-10-06
output:
  distill::distill_article:
    self_contained: false
---

In the past years, I have (co-)authored a couple of R packages that ended up being uploaded to CRAN (`newsanchor`, `pocketapi` and `giedata`). Now, I have always been curious as to whether people actually use them. While one can't know that for sure, one can have a look at the download numbers on CRAN - they are publicly available.

So that is what I did a while, I used a function to download and display the numbers from CRAN, you can find a Gist with a function to do so [here](https://gist.github.com/yannikbuhl/c84c3cbf94cee3091471a2b296a90995#file-cranlogs-r). However, I recently got myself a Raspberry Pi 4 (a Raspberry 2 or 3 will do the job, too) and wanted to have automated CRAN download reports send to me via mail. So I dug in the internet, and here's what I came up with.

First, we need a R script that we can automate that will get us the data from CRAN and save it on the Raspberry Pi's drive so we can send it as a mail attachment:

```{r}
# File name: get-package-data.R

library(cranlogs)
library(dplyr)
library(tibble)

# The download function
cran <- function(package, from = "2019-01-01") {

	mls <- cran_downloads(packages = package, from = from, to = Sys.Date() - 1)

	return(mls)

}

# Read in the "old" data - you will need to create the file once at the start
old_data <- utils::read.csv("cran-hist.csv")

# Count the number of downloads
pocketapi_sum <- cran("pocketapi") %>% summarise(sum = sum(count))
newsanchor_sum <- cran("newsanchor") %>% summarise(sum = sum(count))
giedata_sum <- cran("giedata") %>% summarise(sum = sum(count))

# Calculate the difference to the "old" data
diff_pocket <- pocketapi_sum$sum - old_data$sum[old_data$package == "pocketapi"]
diff_news <- newsanchor_sum$sum - old_data$sum[old_data$package == "newsanchor"]
diff_gie <- giedata_sum$sum - old_data$sum[old_data$package == "giedata"]

# Create a tibble that can be saved to a file
crandata <- tibble(package = c("pocketapi", "newsanchor", "giedata"),
		   date = rep(Sys.time(), 3),
		   sum = c(pocketapi_sum$sum, newsanchor_sum$sum, giedata_sum$sum),
		   diff = c(diff_pocket, diff_news, diff_gie))

# Write the data into a .txt file, our mail attachment
write.table(crandata, "cranlogs.txt", quote = FALSE, row.names = FALSE, 
	    sep = "     |     ")

# Update the file with the historic data
write.csv(crandata, "cran-hist.csv", row.names = FALSE)

```

Now, our mail attachment file looks like this:

```{r}
crandata
```

Next, we use a Python script to send us a mail using the attachment .txt file. I followed the instructions [here](https://www.abstractapi.com/guides/sending-email-with-python). *Be aware that you cannot use Gmail easily anymore because external apps can no longer be allowed to access Gmail. You have to use OAuth2 (see instructions following the link above).*

```{python, eval = FALSE}
# File name: send-mail.py

import smtplib
import ssl
from email.mime.multipart import MIMEMultipart
from email.mime.text import MIMEText
from email.mime.application import MIMEApplication

ctx = ssl.create_default_context()

password = "PASSWORD" # Your password - change!
sender = "sender@mail.com" # Your sender mail address - change!
receiver = "receiver@mail.com" # Receiver address - change!

message = MIMEMultipart("mixed")
message["Subject"] = "Neue CRAN-Downloads"
message["From"] = sender
message["To"] = receiver

plain = """\
        Hello Yannik, 

        there are new CRAN download numbers available.

        Best
        Raspby 4

        """
        
message.attach(MIMEText(plain, "plain"))

# Path to the attachment file - change!
filename = 'cranlogs.txt'

with open(filename, "rb") as f:
    file = MIMEApplication(f.read())

disposition = "attachment; filename=cranlogs.txt"

file.add_header("Content-Disposition", disposition)

message.attach(file)

with smtplib.SMTP_SSL("smtp.gmail.com", port = 465, context = ctx) as server:
    server.login(sender, password)
    server.sendmail(sender, receiver, message.as_string())

```

Finally, just upload these files to your Raspberry Pi and automate them using a Cronjob. In this example, the scripts run every morning and send you an update on your package downloads. Make sure the R script runs before the Python script:

```{bash, eval = FALSE}
30 7 * * * /usr/bin/Rscript --no-restore --no-site-file --no-init-file ~/path/to/your/file/get-package-data.R
35 7 * * * /usr/bin/python ~/path/to/your/file/send-mail.py
```

